Forecasting the 2024 U.S. Presidential Election: Kamala Harris' Projected Victory

### 2024-11-03 paper/paper.qmd
committing code changes to paper.qmd

```
For instance, [@fig-summarystatistics] reveals a total of 107 polls conducted, calculated as the total count of rows in our cleaned dataset. ‘Total polls’ reflects the total number of unique poll entries after filtering for only battleground states, polls with sufficient data quality, and polls that report support percentages for both Trump and Harris. The 107 polls conducted display an average support level of 47.801% for Trump and 47.999% for Harris. This suggests that based on the dataset, Trump has a leading advantage in winning the presidential election race by having a slighty higher average level of support than Harris. The highest percentage of support for Trump in the 107 polls was 51.700% while the lowest recorded support was 43%. For Harris, the highest percentage of support also reached 51.200%, with a minimum of 43.600%. These statistics illustrate the competitive landscape of the presidential election race, without a clear leading candidate in the current race.

@@

## Polling Averages for Trump and Harris Over Time in Battleground States
[@fit-state-polling-average] provides a visual comparison of the current polling support for Donald Trump and Kamala Harris across key battleground states, tracked over time from August 2024 to October 2024. Each line plot represents the average polling percentages for both candidates in a specific state, with red lines denoting Trump’s support and blue lines representing Harris’s support. Across several states, there are noticeably increased fluctuations in polling averages for both candidates closer to the election particularly from late September to October. This pattern may indicate heightened voter interest and engagement as the election date approaches which is November 4th, 2024, it could also reflect changes in public opinion due to campaign events, political debates, or other external factors.
```{r}
#| echo: false
#| eval: true
#| label: fig-state-polling-average
#| fig-cap: "Polling Averages for Trump and Harris Over Time in Battleground States."
#| warning: false

@@

Since it is difficult to conclude a definitive predictive outcome by observing the current data, we utilize our models, described in [@sec-model] to predict the possible outcomes by incorporating the given data. By using the current data, the model can identify the state-specific trends thus generating more accurate predictions of the outcome. Relevant prediction results are provided below.

## State-Level Probability of Trump Winning on Election Day
By incorporating our Bayesian models, we generate the predicted probabilities of winner for both Donald Trump and Kamala Harris in key battleground states, on Election Day (November 5th, 2024). [@fig-state-probabilities] provides a visual summary of these predicted probabilities.
```{r}
#| echo: false
#| eval: true
#| label: fig-state-probabilities
#| fig-cap: "Probability of Trump Winning by State on Election Day."
#| warning: false

@@

## Probability of Winning by State on Election Day 
[@tbl-fig-probabilities] displays the numeric percentage of the probabilities of winning for Trump and Harris. By comparing the predicted probabilities, the predicted winner for each state is assigned accordingly. For instance, the table displays that Trump has a 49.48% predicted probability of winning Arizona compared to 49.24% of Harris’s, making Trump the predicted winner in Arizona. For all six other states, Harris has a higher predicted probability of winning than Trump.
```{r}
#| echo: false
#| eval: true
#| label: tbl-state-probabilities
#| fig-cap: "Probability of Trump Winning by State on Election Day."
#| warning: false

@@

## US Map Showing Predicted Winner by State on Election Day
To visualize the colour distribution of the predicted results, we emloy the US map from [@usmap] to display the results from [@tbl-state-probabilities].

@@

# Print the map
print(us_map)
```
According to [@fig-us-map], Arizona is the only state where Trump has a leading win while all the other six states are led by Harris. Since Harris has a six-to-one lead in the predicted percentage of support for all battleground states, we conclude that Harris is the predicted winner for the 2024 cycle of Presidential Election.

>>>>

```
For instance, @tbl-outcomevar reveals a total of 107 polls conducted, calculated as the total count of rows in our cleaned dataset. ‘Total polls’ reflects the total number of unique poll entries after filtering for only battleground states, polls with sufficient data quality, and polls that report support percentages for both Trump and Harris. The 107 polls conducted display an average support level of 47.801% for Trump and 47.999% for Harris. This suggests that based on the dataset, Trump has a leading advantage in winning the presidential election race by having a slighty higher average level of support than Harris. The highest percentage of support for Trump in the 107 polls was 51.700% while the lowest recorded support was 43%. For Harris, the highest percentage of support also reached 51.200%, with a minimum of 43.600%. These statistics illustrate the competitive landscape of the presidential election race, without a clear leading candidate in the current race.

@@

## Polling Averages for Trump and Harris Over Time in Battleground States
@fig-state-polling-average provides a visual comparison of the current polling support for Donald Trump and Kamala Harris across key battleground states, tracked over time from August 2024 to October 2024. Each line plot represents the average polling percentages for both candidates in a specific state, with red lines denoting Trump’s support and blue lines representing Harris’s support. Across several states, there are noticeably increased fluctuations in polling averages for both candidates closer to the election particularly from late September to October. This pattern may indicate heightened voter interest and engagement as the election date approaches which is November 4th, 2024, it could also reflect changes in public opinion due to campaign events, political debates, or other external factors.
```{r}
#| echo: false
#| eval: true
#| label: fig-state-polling-average
#| fig-cap: "Polling Averages for Trump and Harris Over Time in Battleground States."
#| warning: false

@@

Since it is difficult to conclude a definitive predictive outcome by observing the current data, we utilize our models, described in @sec-model to predict the possible outcomes by incorporating the given data. By using the current data, the model can identify the state-specific trends thus generating more accurate predictions of the outcome. Relevant prediction results are provided below.

## State-Level Probability of Trump Winning on Election Day
By incorporating our Bayesian models, we generate the predicted probabilities of winner for both Donald Trump and Kamala Harris in key battleground states, on Election Day (November 5th, 2024). @fig-state-probabilities provides a visual summary of these predicted probabilities.
```{r}
#| echo: false
#| eval: true
#| label: fig-state-probabilities
#| fig-cap: "Probability of Trump Winning by State on Election Day."
#| warning: false

@@

## Probability of Winning by State on Election Day 
@tbl-state-probabilities displays the numeric percentage of the probabilities of winning for Trump and Harris. By comparing the predicted probabilities, the predicted winner for each state is assigned accordingly. For instance, the table displays that Trump has a 49.48% predicted probability of winning Arizona compared to 49.24% of Harris’s, making Trump the predicted winner in Arizona. For all six other states, Harris has a higher predicted probability of winning than Trump.
```{r}
#| echo: false
#| eval: true
#| label: tbl-state-probabilities
#| fig-cap: "Probability of Trump Winning by State on Election Day."
#| warning: false

@@

## US Map Showing Predicted Winner by State on Election Day
To visualize the colour distribution of the predicted results, we emloy the US map from [@usmap] to display the results from @tbl-state-probabilities.

@@

# Print the map
print(us_map)
```
According to @fig-us-map, Arizona is the only state where Trump has a leading win while all the other six states are led by Harris. Since Harris has a six-to-one lead in the predicted percentage of support for all battleground states, we conclude that Harris is the predicted winner for the 2024 cycle of Presidential Election.
